A Quantitative Genetic Linked Selection Model of the Negative Selection Process in Humans

### 2023-07-23 manuscript/manuscript.tex
ms updates from italy

Given the observed DFE-substitution rate contradiction and that our simulations
indicated that $B(x)$ is more accurately predicted when we use local-rescaling,
we modified our composite-likelihood method to so that it can be run a second
time, on B' maps locally-rescaled by the predicted $\widehat{B}(x)$ from the
initial fit. Intuitively, this is based on the notion that if a neutral allele
experiences an effective population size of $B(x)N$, so too should a selected
allele, and this should be considered in how the SC16 equations are solved.
There is an important caveat about this analysis. Since local-rescaling is done
using the first round of maximum likelihood estimates, there is a possibility
of some circularity, since the $B(x)$ at this position includes the
contribution of the focal segment that is being rescaled. However, the results
do not change based on whether the $B(x)$ is at averaged at the 1 kbp level or
at the megabase scale; for the latter, a single segment makes little
contribution to the average.

With this caveat noted, there are five important but tentative results to draw
from this analysis. First, the estimated mutation rate is implausibly higher,
reaching the upper limit of our optimization boundary of $\mu = 8 \times
10^{-8}$ across all populations in or PhastCons Priority model (Supplementary
Material XXX). When we re-ran optimization with an upper limit of $\mu = 1
\times 10^{-7}$, the mutation rate was again estimated to this (implausibly
large) upper limit, with no considerable changes to model fit or DFE estimates.
Second, all of our leave-one-chromosome-out $R_\text{LOCO}^2$ are about one
percentage point higher than the unrescaled model. This improved fit could be
an artifact of the circularity, so it is unimportant. Third, the DFE estimates
for PhastCons regions is now U-shaped (Supplementary Materials Figure
\ref{suppfig:rescale-dfe}), with $75-77\%$ of mass being placed on a weakly
deleterious class, $s=10{^-5}$. Interestingly, this is the first of all of our
models where such an appreciable mass has been placed on a midpoint in our
grid; in all other cases, non-strongly estimates were neutral ($s=10^{-8}$).
Fourth, predicted pairwise diversity is nearly identical to our original,
non-locally-rescaled fit (see Supplementary Materials Figure
\ref{suppfig:rescale}). Finally, since local-rescaling increases the DFE mass
over $s < 10^{-3}$, mutations in PhastCons regions now have the possibility of
fixation. We find that local-rescaling the PhastCons Priority, leads the
predicted substitution rates in PhastCons regions to be much closer to observed
levels (blue line, Figure \ref{fig:figure-6}).

Finally, as a robustness check, we investigated the extent to which
local-rescaling differed from the B' maps across selection parameters. We find
minor differences between the locally-rescaled and standard B' maps in the $0.2
\le 2N_e s \le 2$ range; the maps were identical under strong selection ($2N_e
s= 20$) as expected (Supplementary Materials Figure \ref{suppfig:rescale-b}).
Overall, this suggests that models that fit the linked selection are worryingly
sensitive the theoretic B' values in the $2Ns \le 1$ domain. Local-rescaling
turns out to only slightly alter the B' maps, yet significantly modifies the
DFE estimates. This brings the deleterious substitution rate in agreement with
observations (since this is predicted with a fixed $\mu = 1.5 \times 10^{-8}$;
however, the maximum likelihood estimate of mutation is implausibly high.

@@

We hypothesized that this could be because in this domain, a segment under
selective constraint experiences a local draft-effective population size of
$B(x)N$, and not just $N$. This local draft-effective population size is
induced by selection at other segments that is not being taken into account by
classic BGS theory or our standard model. When we experimented by fitting our
model and then using the predicted reduction map to local-rescale $N_e$ to the
draft-effective population size $N_d = \widehat{B}(x)$, we find the
disagreement between predicted and observed substitution rates disappears. This
is expected, since locally-rescaled fits result in DFE estimates to have an
appreciable mass on weakly deleterious selection coefficients. While this
brings substitution rates into agreement, we encounter the same problem found
by \textcite{McVicker2009-ax}: the estimated mutation rate is too high. We
emphasize that although there are several caveats with the local-rescaling
approach, in practice it is altering the B' maps only slightly in the domain of
$2N_e s \approx 1$. Even if the local-rescaling approximation to selective
interference is incorrect, these results suggest parameter estimates are
extremely sensitivity to how accurate the B' maps are in this domain.

Unfortunately, this work opens up more questions about the nature of selection
in humans that it answers. The implausibly high estimate of mutation rates
could suggest that negative selection is not the only source of fitness
variance generating the genome-wide linked selection signal in humans. Positive
selection on polygenic variation along the genome could be another source of
fitness variance contributing to the weak selection signal. Alternatively,
negative selection could be the main source of fitness variance, but the
complexity of selective interference may be poorly approximated by the
local-rescaling approach. Overall, understanding how the sources of human
fitness variance shape the striking patterns of diversity along the human
genome still remains an open riddle.

>>>>

Given the observed DFE-substitution rate contradiction and that our simulations
indicated that $B(x)$ is more accurately predicted when we use local-rescaling,
we modified our composite-likelihood method to so that it can be run a second
time, on B' maps locally-rescaled by the predicted $\widehat{B}(x)$ from the
initial fit. Intuitively, this is based on the notion that if a neutral allele
experiences an effective population size of $B(x)N$, so too should a selected
allele, and this should be considered in how the SC16 equations are solved.
There is an important caveat about this analysis.

There are five important but tentative results to draw from this analysis.
First, the estimated mutation rate is implausibly higher, reaching the upper
limit of our optimization boundary of $\mu = 8 \times 10^{-8}$ across all
populations in or PhastCons Priority model (Supplementary Material XXX). When
we re-ran optimization with an upper limit of $\mu = 1 \times 10^{-7}$, the
mutation rate was again estimated to this (implausibly large) upper limit, with
no considerable changes to model fit or DFE estimates. Second, all of our
leave-one-chromosome-out $R_\text{LOCO}^2$ are about one percentage point
higher than the unrescaled model. This improved fit could be an artifact of the
circularity, so it is unimportant. Third, the DFE estimates for PhastCons
regions is now U-shaped (Supplementary Materials Figure
\ref{suppfig:rescale-dfe}), with $75-77\%$ of mass being placed on a weakly
deleterious class, $s=10{^-5}$. Interestingly, this is the first of all of our
models where such an appreciable mass has been placed on a midpoint in our
grid; in all other cases, non-strongly estimates were neutral ($s=10^{-8}$).
Fourth, predicted pairwise diversity is nearly identical to our original,
non-locally-rescaled fit (see Supplementary Materials Figure
\ref{suppfig:rescale}). Finally, since local-rescaling increases the DFE mass
over $s < 10^{-3}$, mutations in PhastCons regions now have the possibility of
fixation. We find that local-rescaling the PhastCons Priority, leads the
predicted substitution rates in PhastCons regions to be much closer to observed
levels (blue line, Figure \ref{fig:figure-6}).

Finally, we one important caveat about this analysis. Since local-rescaling is
done using the first round of maximum likelihood estimates, there is a
possibility of some circularity, since the $B(x)$ at this position includes the
contribution of the focal segment that is being rescaled. However, two
observations indicate this would matter little. First, the results do not
change based on whether the $B(x)$ is at averaged at the 1 kbp level or at the
megabase scale; for the latter, a single segment makes little contribution to
the average. Second, we investigated the extent to which local-rescaling
differed from the B' maps across selection parameters. We find minor
differences between the locally-rescaled and standard B' maps in the
nearly-neutral domain ($0.2 \le 2N_e s \le 2$); the maps were identical under
strong selection ($2N_e s= 20$) as expected (Supplementary Materials Figure
\ref{suppfig:rescale-b}). Moreover, the correlations between the standard and
local-rescaled B' maps across the genome are high (100\% for $2N_e s = 20$,
96.5\% for $2N_e s = 2$, and 60.21\% for $2N_e s = 0.2$). The overall realized
effect of local-rescaling is to just alter how deep the ``U" is in the
relationship between the reduction factor $B$ and the selection coefficient
(Supplementary Materials Figure \ref{suppfig:rescale-effects}).

Overall, this suggests that models that fit the linked selection are worryingly
sensitive the theoretic B' values in the $2Ns \le 1$ domain. Local-rescaling
turns out to only slightly alter the B' maps, yet significantly modifies the
DFE estimates. This brings the deleterious substitution rate in agreement with
observations (since this is predicted with a fixed $\mu = 1.5 \times 10^{-8}$;
however, the maximum likelihood estimate of mutation is implausibly high.

@@

We hypothesized that this could be because in this domain, a segment under
selective constraint experiences a local draft-effective population size of
$B(x)N$, and not just $N$. This local draft-effective population size is
induced by selection at other segments that is not being taken into account by
classic BGS theory or our standard model. When we experimented by fitting our
model and then using the predicted reduction map to local-rescale $N_e$ to the
draft-effective population size $N_d = \widehat{B}(x)$, we find the
disagreement between predicted and observed substitution rates disappears. This
is expected, since locally-rescaled fits result in DFE estimates to have an
appreciable mass on weakly deleterious selection coefficients. While this
brings substitution rates into agreement, we encounter a similar problem found
by \textcite{McVicker2009-ax}: the estimated mutation rate is too high. We
emphasize that although there are several caveats with the local-rescaling
approach, in practice it is altering the B' maps only slightly in the domain of
$2N_e s \approx 1$ (deepening the bottom of the ``U"). Even if the local-rescaling
approximation to selective interference is incorrect, these results suggest
parameter estimates are extremely sensitivity to how accurate the B' maps are
in this domain.

Unfortunately, this work opens up more questions about the nature of selection
in humans that it answers. The implausibly high estimate of mutation rates
could suggest that negative selection is not the only source of fitness
variance generating the genome-wide linked selection signal in humans. Positive
selection on polygenic variation along the genome could be another source of
fitness variance contributing to the weak selection signal. Alternatively,
negative selection could be the main source of fitness variance, but the
complexity of selective interference may be poorly approximated by the
local-rescaling approach. Another possibility is that our assumption throughout
of additive effects may lead to biases given deleterious mutations likely
recessive effects. Overall, understanding how the sources of human fitness
variance shape the striking patterns of diversity along the human genome still
remains an open riddle.
